Extreme Value Statistics Draft 1

### 2022-03-21 Draft#1.Rmd
draft2

>>>>

Modeling extremal dependence

We first construct the pseudo-polar coordinates and then extract the angles w corresponding to extreme events.

```{r}
r.obs=apply(data.stand,1,sum)
w.obs=data.stand/r.obs
r_0=quantile(r.obs,0.95)
w.H=w.obs[r.obs>r_0,]
```

### 3.2) Estimation of joint exceedances above the overall 95% quantile

```{r}
q95=quantile(mydata,0.95)
q95
prob.component=apply(mydata<q95,2,mean)
u.star=1/(1-prob.component)
u.star
ell.min=function(w,u){min(w/u)} #already defined in preceding exercice
d=5
n_0=nrow(w.H)
#estimate the probability of the joint exceedance event:
d/n_0*sum(apply(w.H,1,ell.min,u=u.star))
#compare to observed empirical probability:
mean(apply(t(t(data.stand)/u.star)>1,1,all))

Estimation of a parametric Dirichlet model for the angular measure

```{r}
library(MCMCpack) #for Dirichlet density 
library(maxLik) #for numerical maximisation of the log-likelihood
```

#define the log-likelihood function:
llfun=function(alpha){
if(alpha<=0) return(-Inf) #inadmissible parameters
 sum(log(ddirichlet(w.H,alpha=rep(alpha,d))))
}

initpar=1 # define the initial value of the parameter
fit=maxLik(llfun,start=initpar)
summary(fit)
